Make any stupid plot you want

### 2021-03-01 draft/sections/intro.tex
rework of contribution section

\begin{figure}[H]
    \includegraphics[width=1\textwidth]{figures/intro/dataset_diagram.png}
    \caption{One way to describe data is by the connectivity of the points in the dataset. A database for example is often discrete unconnected points, while an image is an implicitely connected 2D grid. This image is from the Data Representation chapter of the MayaVi 4.7.2 documentation.\cite{DataRepresentationMayavi}}
    \label{fig:intro_data_format}
\end{figure}
As shown in figure~\ref{fig:intro_data_format}, there are many distinct ways of encoding each specific type of structure, while as mentioned in section~\ref{sec:intro_data_tools} APIs are clearer when structured around a common data representation. Fiber bundles were proposed by Butler as one such representation because they encode the continuity of the data separately from the types of variables and are flexible enough to support discrete and ND continuous datasets \cite{butlerVisualizationModelBased1989,butlerVectorBundleClassesForm1992}. Since Butler's model lacks a robust way of describing variables, we fold in Spivak's Simplacial formulation of databases \cite{spivakDatabasesAreCategories2010,spivakSIMPLICIALDATABASES} so that we can encode a schema like description of the data in the fiber bundle. In this work we will refer to the points of the dataset as \textit{records} to indicate that a point can be a vector of heterogenous elements. Each \textit{component} of the record is a single object, such as a temperature measurement, a color value, or an image. We also generalize \textit{component} to mean all objects in the dataset of a given type, such as all temperatures or colors or images. The way in which these records are connected is the \textit{connectivity}, \textit{continuity}, or more generally \textit{topology}.

@@

Often this topology has metadata associated with it, describing for example when or where the measurement was taken. Building on the idea of metadata as \textit{keys} and their associated \textit{value} proposed by Munzner \cite{munznerChDataAbstraction}, we propose that information rich metadata are part of the components and instead the values are keyed on coordinate free structural ids. In contrast to Munzner's model where the semantic meaning of the key is tightly coupled to the position of the value in the dataset, our model allows for renaming all the metadata, for example changing the coordinate systems or time resolution, without imposing new semantics on the underlying structure.

@@

One example of highly expressive visualizations are the data portraits by Du Bois shown in figure~\ref{fig:intro_dubois}. While the Du Bois charts are different from the usual scatter, line, and plot charts, they conform to the constraint that a graphic is a structure preserving map from data to visual representation. Figure~\ref{fig:intro_dpa} is semantically similar to a bar chart in that the lengths of the segments are mapped to the values, but in this chart the segments are stacked together. Figure~\ref{fig:intro_dpb} is a multicolored line chart where the color shifts are at periods of political significance. In figure~\ref{fig:intro_dbc}, Du Bois combines a graphical representation where glyph size varies by population with a figurative representation of those glyphs as the countries the data is from, which means that the semantic and numerical properties of the data are preserved in the graph. Figure~\ref{fig:intro_dpb} is simply a treemap\cite{heerTourVisualizationZoo2010} with space between the marks. Since the Du Bois data portraits meet the criteria of a faithful visual representation, we propose a mathematical framework and implementation that allows us to express the Du Bois charts and common chart types with equal fidelity.

>>>>

\begin{figure}[H]
    \includegraphics[width=1\textwidth]{figures/intro/dataset_diagram.png}
    \caption{One way to describe data is by the connectivity of the points in the dataset. A database for example is often discrete unconnected points, while an image is an implicitely connected 2D grid. This image is from the Data Representation chapter of the MayaVi 4.7.2 documentation.\cite{DataRepresentationMayavi}}
    \label{fig:intro_data_format}
\end{figure}
As shown in figure~\ref{fig:intro_data_format}, there are many distinct ways of encoding each specific type of structure, while as mentioned in section~\ref{sec:intro_data_tools} APIs are clearer when structured around a common data representation. Fiber bundles were proposed by Butler as one such representation because they encode the continuity of the data separately from the types of variables and are flexible enough to support discrete and ND continuous datasets \cite{butlerVisualizationModelBased1989,butlerVectorBundleClassesForm1992}. Since Butler's model lacks a robust way of describing variables, we fold in Spivak's Simplicial formulation of databases \cite{spivakDatabasesAreCategories2010,spivakSIMPLICIALDATABASES} so that we can encode a schema like description of the data in the fiber bundle. In this work we will refer to the points of the dataset as \textit{records} to indicate that a point can be a vector of heterogenous elements. Each \textit{component} of the record is a single object, such as a temperature measurement, a color value, or an image. We also generalize \textit{component} to mean all objects in the dataset of a given type, such as all temperatures or colors or images. The way in which these records are connected is the \textit{connectivity}, \textit{continuity}, or more generally \textit{topology}.

@@

Often this topology has metadata associated with it, describing for example when or where the measurement was taken. Building on the idea of metadata as \textit{keys} and their associated \textit{values} proposed by Munzner \cite{munznerChDataAbstraction}, we propose that information rich metadata are part of the components and instead the values are keyed on coordinate free structural ids. In contrast to Munzner's model where the semantic meaning of the key is tightly coupled to the position of the value in the dataset, our model considers keys to be a pure reference to topology. This allows the metadata to be altered, for example by changing the coordinate systems or time resolution, without imposing new semantics on the underlying structure.

@@

One example of highly expressive visualizations are the data portraits by Du Bois shown in figure~\ref{fig:intro_dubois}. While the Du Bois charts are different from the usual scatter, line, and plot charts, they conform to the constraint that a graphic is a structure preserving map from data to visual representation. Figure~\ref{fig:intro_dpa} is semantically similar to a bar chart in that the lengths of the segments are mapped to the values, but in this chart the segments are stacked together. Figure~\ref{fig:intro_dpb} is a multicolored line chart where the color shifts are at periods of political significance. In figure~\ref{fig:intro_dbc}, Du Bois combines a graphical representation where glyph size varies by population with a figurative representation of those glyphs as the countries the data is from, which means that the semantic and numerical properties of the data are preserved in the graph. Figure~\ref{fig:intro_dpd} is simply a treemap\cite{heerTourVisualizationZoo2010} with space between the marks. Since the Du Bois data portraits meet the criteria of a faithful visual representation, we propose a mathematical framework and implementation that allows us to express the Du Bois charts and common chart types with equal fidelity.
